Convolutional neural networks performing a visual search task show attention-like limits on accuracy when trained to generalize across multiple search stimuli

### 2019-06-05 docs/ccneuro2019/paper.tex
revamp first section of intro on visual search

\section{Introduction}
\subsection{Attention}
Multiple converging lines of evidence seem to suggest that training neural networks to 
perform tasks with human-like abilities causes them to acquire brain-like representations.

- If brain-like representations are both necessary and sufficient for human-like 
intelligence, then shouldn't the converse also be true? Shouldn't neural networks that 
have acquired brain-like representations behave like humans when asked to perform related 
tasks? We test this

In these fields, theories of attention mechanisms are often built on 
results produced by experimental manipulations that impair behavior.
Perhaps the pre-eminent example of this approach to understanding attention is a 
a long-established form of visual search task [@wolfeVisualSearch1998]. In this behavioral
paradigm, subjects view a set of discrete items on a flat background and report whether a 
target is present or absent among distractors. Classically, results from this paradigm 
have been used to support different theories of \bf{selective covert visual attention}, 
a term that requires some defintion. The term "selective" is simply supposed to mean 
some mechanism that selects some visual inputs for further processing while necessarily 
discarding others. The word "covert" in this context means any mechanism that occurs 
"centrally", i.e. inside the brain, without overt orientation, i.e. without eye or head 
movements.
- here we do not

Our goal here is not to resolve the debate about whether selective covert visual 
attention is serial or parallel. Rather our goal is to better understand attention as 
defined in cognitive science and reconcile that with the definition used in artificial 
intelligence.

\subsection{visual search}
\textbf{set size effects}

>>>>

\section{Introduction}
\subsection{What are the factors that limit visual search?}
What limits our ability to find what we are looking for in the cluttered noisy world we 
see around us? One of the principle tasks that has been used to investigate this question
is a classic form of visual search task, in which subjects view a stimulus on each trial, 
made up of discrete sets of items on a solid-colored background, and report whether a 
target is present among distractors [@wolfeVisualSearch1998]. 
Most studies experimentally manipulate factors such as features of the targets and 
distractors, in order to identify those factors limit visual search 
[@ecksteinVisualSearchRetrospective2011]. Importantly, these factors that *limit* visual 
search performance form the basis of seminal theories of selective visual attention. 
As we review briefly below for context, there are essentially two families of selective 
attention mechanisms. The results supporting both mechanisms remain unreconciled, in spite
of nearly half a century of studies based on this task. The core question is: to what 
extent can the limitations on visual search be attributed to some computation that is 
required to bind features into items, and to what extent can those limitations be 
explained by other aspects of computation, such as noisy internal representations and 
decision-making criteria? Another way of stating this problem is to ask: if there 
was some algorithm that could produce an ideal observer capable of maximally utilizing 
the available information to carry out the task, would that ideal observer still be 
subject to some ceiling of performance, due simply to the constraints of the task?

For context, we briefly review the two competing theoretical mechanisms of selective 
attention, and describe how they depend in part on how the visual search task is 
performed. The first family consists of serial mechanisms. One of the most influential 
early incarnations is known as Feature Integration Theory (FIT) 
[@treismanFeatureintegrationAttention1980]. Essentially, FIT posits that the visual system
can process single "features", such as color or orientation, in parallel, but when 
required to utilize multiple features, it uses some sort of serial computation, i.e. one 
that takes time. Evidence for this theory came from visual search experiments where 
subjects were shown stimuli until responding, and their reaction time was measured. 
Plotting reaction time as a function of set size revealed 
lines whose slope were near zero when a single feature distinguished the target from 
distractors (e.g., "red" v. "green"). The slope of reaction time v. set size increased 
sharply, though, when distinguishing the target from distractors required finding a 
conjunction of features (such as "red vertical" v. "red horizontal or green vertical"). 
However, subsequent studies showed that slopes could not be cleanly segregated in a way 
that supported distinct parallel and serial mechanisms [@wolfeWhatCanMillion1998]. This 
led to a more general model of visual search which seeks to account for results from 
multiple tasks, but still features at its core a serial processing step
[@wolfeGuidedSearchRevised1994].
The second family of search mechanisms are parallel, and arose in large part in reaction 
to the way the visual search task was carried out when investigating
serial mechanisms. Early studies that measured reaction times left several factors 
uncontrolled, such as target-distractor similarity [@duncanVisualSearchStimulus1989], 
drops in acuity outside the fovea, eye movements, and effects resulting from visual 
crowding [@bergenRapidDiscriminationVisual1983; @ecksteinLowerVisualSearch1998]. 
Hence researchers designed versions of the visual search task that controlled for such 
factors[@palmerPsychophysicsVisualSearch2000]. Crucially, they showed subjects the 
stimulus only briefly, to prevent eye movements, and measured accuracy instead of reaction
time. These researchers argued that set size-dependent decreases in accuracy could be 
explained by totally parallel mechanisms of attention, where performance was impaired by 
noise. Computational models of parallel mechanisms succesfully explained results from 
single feature and conjunction search stimuli, in terms of distributions and 
discrimination thresholds borrowed from signal detection theory that psychophysicists 
often employ.
